### 2021-12-03 text/outline/draft.tex
update real data plotting code

\begin{document}
\noindent
Tim, Gene, Kathryn
\begin{center} 
\textbf{A class of exponential family measurement error models applied to CRISPR genome editing and single-cell sequencing}
\end{center}

\begin{abstract}
CRISPR genome engineering and single-cell sequencing have transformed biological discovery. Single-cell CRISPR screens unite these two technologies, linking genetic perturbations in individual cells to changes in gene expression and illuminating regulatory networks underlying diseases. Despite their promise, single-cell CRISPR screens present substantial statistical challenges. We demonstrate on real data that a standard method for estimation and inference in single-cell CRISPR screens — “thresholded regression” — exhibits attenuation bias and a bias-variance tradeoff as a function of an intrinsic tuning parameter. We recover these phenomena in precise theoretical terms in an idealized Gaussian setting. Next, we introduce GLM-EIV (``GLM-based errors-in-variables''), a new method for single-cell CRISPR screen analysis. GLM-EIV generalizes the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed, overcoming limitations of thresholded regression. We develop a computational infrastructure to deploy GLM-EIV across tens or hundreds of nodes on clouds (e.g., Microsoft Azure) and high-performance clusters. Leveraging this infrastructure, we apply GLM-EIV to analyze two recent, large-scale, single-cell CRISPR screen datasets, yielding new biological insights.
\end{abstract}
\tableofcontents
\newpage
\section{Introduction}
CRISPR is a genome engineering tool that has enabled scientists to precisely edit human and nonhuman genomes, opening the door to new medical therapies \cite{Rothgangl2021,Musunuru2021} and transforming basic biology research \cite{Przybyla2021}. Recently, scientists have paired CRISPR genome engineering with single-cell sequencing \cite{Dixit2016,Datlinger2017}. The resulting assays, known as a ``single-cell CRISPR screens,'' link genetic perturbations in individual cells to changes in gene expression, illuminating regulatory networks underlying human diseases and other traits \cite{Morris2021a}.

@@

\subsection{Computation}
We develop a computational infrastructure to apply GLM-EIV to large-scale, single-cell CRISPR screen data.\footnote{Pipeline available at \texttt{github.com/timothy-barry/glmeiv-pipeline}.} The infrastructure leverages Nextflow, a programming language that facilitates building data-intensive pipelines \cite{DITommaso2017}, and ondisc\footnote{Package available at \texttt{github.com/timothy-barry/ondisc}; preprint forthcoming.}, an R package that we developed (in a separate project) to facilitate large-scale computing on single-cell data. Nextflow and ondisc together enable the construction of highly portable single-cell pipelines: one can analyze data \textit{out-of-memory} on a laptop or in a \textit{distributed} fashion across tens or hundreds of nodes on a cloud (e.g., Microsoft Azure, Google Cloud) or high-performance cluster.

@@

First, we observed that GLM-EIV dominated thresholded regression on all statistical metrics: GLM-EIV exhibited lower bias (row 1) and mean squared error (row 2) than thresholded regression; additionally, GLM-EIV had superior confidence interval coverage (row 3) despite having produced generally narrower confidence intervals (row 4). Intuitively, GLM-EIV outperformed the thresholding method because (i) GLM-EIV leveraged information from \textit{both} modalities (rather than the gRNA modality alone) to assign perturbation identities to cells, and (ii) GLM-EIV produced soft rather than hard assignments, capturing the inherent uncertainty in whether a perturbation occurred. Next, we found that accelerated GLM-EIV performed as well as ``vanilla'' GLM-EIV on the statistical metrics (rows 1-4) while substantially reducing computational cost (bottom row). In fact, the execution time of accelerated GLM-EIV was approximately within an order of magnitude of that of the thresholding method (bottom row).

@@

>>>>

\begin{document}
\noindent
Tim, Gene, Kathryn
\begin{center} 
\textbf{A class of exponential family measurement error models, with application to CRISPR genome editing and single-cell sequencing}
\end{center}

\begin{abstract}
CRISPR genome engineering and single-cell sequencing have transformed biological discovery. Single-cell CRISPR screens unite these two technologies, linking genetic perturbations in individual cells to changes in gene expression and illuminating regulatory networks underlying diseases. Despite their promise, single-cell CRISPR screens present substantial statistical challenges. We demonstrate on real data that a standard method for estimation and inference in single-cell CRISPR screens — “thresholded regression” — exhibits attenuation bias and a bias-variance tradeoff as a function of an intrinsic tuning parameter. We recover these phenomena in precise theoretical terms in an idealized Gaussian setting. Next, we introduce GLM-EIV (``GLM-based errors-in-variables''), a new method for single-cell CRISPR screen analysis. GLM-EIV generalizes the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed, overcoming limitations of thresholded regression in challenging problem settings. We develop a computational infrastructure to deploy GLM-EIV across tens or hundreds of nodes on clouds (e.g., Microsoft Azure) and high-performance clusters. Leveraging this infrastructure, we apply GLM-EIV to analyze two recent, large-scale, single-cell CRISPR screen datasets, yielding new biological insights.
\end{abstract}
\tableofcontents
\newpage
\section{Introduction}
CRISPR is a genome engineering tool that has enabled scientists to precisely edit human and nonhuman genomes, opening the door to new medical therapies \cite{Rothgangl2021,Musunuru2021} and transforming basic biology research \cite{Przybyla2021}. Recently, scientists have paired CRISPR genome engineering with single-cell sequencing \cite{Dixit2016,Datlinger2017}. The resulting assays, known as a ``single-cell CRISPR screens,'' link genetic perturbations in individual cells to changes in gene expression, illuminating regulatory networks underlying human diseases and other traits \cite{Morris2021a}.

@@

\subsection{Computing}
We developed a computational infrastructure to apply GLM-EIV to large-scale, single-cell CRISPR screen data.\footnote{Pipeline available at \texttt{github.com/timothy-barry/glmeiv-pipeline}.} The infrastructure leverages Nextflow, a programming language that facilitates building data-intensive pipelines \cite{DITommaso2017}, and ondisc\footnote{Package available at \texttt{github.com/timothy-barry/ondisc}; preprint forthcoming.}, an R package that we developed (in a separate project) to facilitate large-scale computing on single-cell data. Nextflow and ondisc together enable the construction of highly portable single-cell pipelines: one can analyze data \textit{out-of-memory} on a laptop or in a \textit{distributed} fashion across tens or hundreds of nodes on a cloud (e.g., Microsoft Azure, Google Cloud) or high-performance cluster.

@@

First, we observed that GLM-EIV dominated thresholded regression on all statistical metrics: GLM-EIV exhibited lower bias (row 1) and mean squared error (row 2) than thresholded regression; additionally, GLM-EIV had superior confidence interval coverage (row 3) despite having produced generally narrower confidence intervals (row 4). Intuitively, GLM-EIV outperformed the thresholding method because (i) GLM-EIV leveraged information from \textit{both} modalities (rather than the gRNA modality alone) to assign perturbation identities to cells, and (ii) GLM-EIV produced soft rather than hard assignments, capturing the inherent uncertainty in whether a perturbation occurred. Next, we found that accelerated GLM-EIV performed as well as ``vanilla'' GLM-EIV on the statistical metrics (rows 1-4) despite have substantially lower computational cost (bottom row). In fact, the execution time of accelerated GLM-EIV was approximately within an order of magnitude of that of the thresholding method (bottom row).

@@

We see the practical use of this method as follows. First, deploy GLM-EIV to some random subset of pairs to assess the extent of background contamination. If the background contamination sufficiently high, apply GLM-EIV to analyze the data directly. On the other hand, if the background contamination is low, the thresholding method is an appealing method due to its speed and stability. A challenge, as we have shown, is that the selected threshold exerts a substantial impact on the results. GLM-EIV approximates the Bayes-optimal decision boundary, which, as we showed in an idealized Gaussian setting, minimizes attenuation bias when the gRNA count distribution is sufficiently-well separated. Therefore, in problem settings where thresholding is tenable, GLM-EIV provides a statistically principled, scientifically reproducible, empirically effective threshold selection procedure.

To our knowledge, this is the first statistical work to study single-cell CRISPR screens. One of our objectives is to bring to the attention of the statistical community an emerging class of functional genomics assays that likely will exert a major impact on biology research. Additionally, this is the first work to leverage the ondisc-Nextflow-cloud/HPC technology stack, a powerful and user-friendly set of tools for large-scale single-cell analysis that other researchers (especially statisticians) likely will find useful.
